Von Neumann-Morgenstern and Savage Theorems for Causal Decision Making

### 2020-08-29 Theory_and_decision/Draft/Theory_and_decision_draft.tex
añadiendo citas

\subsection{Some Theory}
\begin{definition}
An uncertain environment is the tuple $(\Omega, \mathcal{A},\mathcal{C},\mathcal{E})$. Where $\mathcal{A}$ is a non-empty set of available actions, $\mathcal{C}$ a set of consequences and $\mathcal{E}$ an algebra of events over $\Omega$. 
\end{definition}

@@

\subsection{Causal Graphical Models}
A Causal Graphical Model (CGM) consists of a set of random variables $\mathcal{X}=\{ X_1,...,X_n \}$, and a Directed Acyclic Graph (DAG) whose nodes are in correspondence with the variables in $\mathcal{X}$ and whose edges represent relations of cause-effect. Also, the model, which up to this point is nothing more than a Bayesian Network with causal semantics, is enriched with an operator named $do()$ which is defined over graphs and whose action is described as follows: given $\mathbf{X} \subseteq \mathcal{X}$ and $\mathbf{x} = \{ x_{i_1}, x_{i_2}, ... , x_{i_j} \} \in Val(\mathcal{X})$ the action $do(\mathbf{X} = \mathbf{x} )$ corresponds to assign to each $X_j \in \mathbf{X}$ the value $x_{x_{i_j}}$ and to delete every incoming edges into the node corresponding to each $X_j$ in the graph $\mathcal{G}$ \citep{pearl2009causality,koller2009probabilistic,sucar2015probabilistic}. To apply the $do()$ operator over a variable (or set of variables) is also called as an \textit{intervention} over the variable. It is this interventional operator which separates associative models from causal models, since it provides the solution of the “probability raising” attempts to define causality, since if $X$ is a cause for $Y$ then $P(Y | do(X)) > P(Y)$ \citep{pearl2018why}. In fact, using the \textit{Do} operator we give a definition of a variable $X$ causing $Y$.
\begin{definition}
Let $\mathcal{G}$ a Causal Graphical Model and $P_\mathcal{G}$ its corresponding distribution and let $X$ and $Y$ variables in the model. We say that $X$ \textit{causes} $Y$ if 
\[ P_\mathcal{G}(Y | do(X=x) \neq P_\mathcal{G}(Y | do(X=x'). \]
\end{definition}
It is required that the probability distribution that results from an intervention over a variable is Markov compatible with the graph; this is, the resulting interventional distribution is equivalent to the product of the conditional probability of every variable given its parents in the intervened graph \citep{sucar2015probabilistic}.
\subsubsection{The identifiability problem}
Under what conditions can causal inquiries be answered in terms of purely \textit{observational} data? It is known that if the Markov Causal condition and Causal Faithfulness hold, then it is possible to identify the causal graph up to related variables without direction (Markov equivalence) \citep{peters2012identifiability,mooij2016distinguishing}.

>>>>

\subsection{Preliminary Definitions}
\begin{definition}
An uncertain environment is the tuple $(\Omega, \mathcal{A},\mathcal{C},\mathcal{E})$. Where $\mathcal{A}$ is a non-empty set of available actions, $\mathcal{C}$ a set of consequences and $\mathcal{E}$ an algebra of events over $\Omega$. 
\end{definition}

@@

\subsection{Causal Graphical Models}
A Causal Graphical Model (CGM) \citep{koller2009probabilistic,sucar2015probabilistic} consists of a set of random variables $\mathcal{X}=\{ X_1,...,X_n \}$, and a Directed Acyclic Graph (DAG) whose nodes are in correspondence with the variables in $\mathcal{X}$ and whose edges represent relations of cause-effect. Also, the model, which up to this point is nothing more than a Bayesian Network with causal semantics, is enriched with an operator named $do()$ which is defined over graphs and whose action is described as follows: given $\mathbf{X} \subseteq \mathcal{X}$ and $\mathbf{x} = \{ x_{i_1}, x_{i_2}, ... , x_{i_j} \} \in Val(\mathcal{X})$ the action $do(\mathbf{X} = \mathbf{x} )$ corresponds to assign to each $X_j \in \mathbf{X}$ the value $x_{x_{i_j}}$ and to delete every incoming edges into the node corresponding to each $X_j$ in the graph $\mathcal{G}$ \citep{pearl2009causality,koller2009probabilistic,sucar2015probabilistic}. To apply the $do()$ operator over a variable (or set of variables) is also called as an \textit{intervention} over the variable. It is this interventional operator which separates associative models from causal models, since it provides the solution of the “probability raising” attempts to define causality, since if $X$ is a cause for $Y$ then $P(Y | do(X)) > P(Y)$ \citep{pearl2018why}. In fact, using the \textit{Do} operator we give a definition of a variable $X$ causing $Y$.
\begin{definition}
Let $\mathcal{G}$ a Causal Graphical Model and $P_\mathcal{G}$ its corresponding distribution and let $X$ and $Y$ variables in the model. We say that $X$ \textit{causes} $Y$ if 
\[ P_\mathcal{G}(Y | do(X=x)) \neq P_\mathcal{G}(Y | do(X=x')). \]
\end{definition}
It is required that the probability distribution that results from an intervention over a variable is Markov compatible with the graph; this is, the resulting interventional distribution is equivalent to the product of the conditional probability of every variable given its parents in the intervened graph \citep{sucar2015probabilistic}.
\subsubsection{The identifiability problem}
Under what conditions can causal inquiries be answered in terms of purely \textit{observational} data? It is known that if the Markov Causal condition and Causal Faithfulness hold, then it is possible to identify the causal graph up to related variables without direction (Markov equivalence) \citep{peters2012identifiability,mooij2016distinguishing}.
